Pushdown Control-Flow Analysis via Refunctionalization (Functional Pearl)

### 2018-03-16 paper.tex
update

As we mentioned previously, the components of states are lifted as arguments to
@aeval@:
\begin{lstlisting}
def aeval(e: Expr, env: Env, store: BStore, time: Time, continue: Cont): Ans = ......
type Cont = Ans => Ans
\end{lstlisting}

@@

\textbf{First Version of Evaluation Function.}
Since our first version of @aeval@ just implements the non-deterministic
evaluation, and given the defined @nd@, the evaluation function can be easily
defined.
The outer call to function @nd@ evaluates over a set of possible closures of
@f@. For each closures, we go into its body with new environment and store.
Remember that function @aeval@ returns a set of @VS@ to its continuation,
so the @bodyvss@ is a set of @VS@ that represents values from possible
computation pathes when evaluating @body@.

Then for the inner call of function @nd@ evaluates over the set of body values.
For each @VS@, we use the time and store
The inner application of @aeval@ returns values of one computation path to its
continuation, where the continuation will accumulate the result and transfer
evaluation to the next body value.

\begin{lstlisting}
def aeval(e: Expr, env: Env, store: BStore, time: Time, continue: Cont): Ans = {
  val new_time = (e::time).take(k)
  e match {
    case Let(x, App(f, ae), e) =>
      val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
      nd[Clos, Ans](closures, Set[VS](), { case (clos, accouter, closnd) =>
        val Clos(Lam(v, body), c_env) = clos
        val baddr = allocBind(v, new_time)
        val new_env = c_env + (v -> baddr)
        val new_store = store.update(baddr, atomicEval(ae, env, store)) 
        aeval(body, new_env, new_store, new_time, (bodyvss: Set[VS]) => {
          nd[VS, Ans](bodyvss, Set[VS](), { case (vs, accinner, bdnd) => 
            val VS(vals, time, store) = vs
            val new_env = env + (x -> baddr)
            val new_store = store.update(baddr, vals)
            aeval(e, new_env, new_store, time, 
                  /* accumulate the values of one path, and call next body value*/
                  (evss: Ans) => bdnd(accinner ++ evss)) 
          },
          /* accumulate the values of pathes to accouter, and call next closure */
          (evss: Ans) => closnd(evss ++ accouter)) 
        })
      }, 
      continue)
    case ae if isAtomic(ae) => 
      val ds = atomicEval(ae, env, store)
      continue(Set(VS(ds, new_time, store))) 
  } 
}
\end{lstlisting}

@@

Pushdown control-flow is a property in analysis that precisely models
the runtime call-stack of the analyzed program. 
A pushdown control-flow analysis provides as exact as runtime return-flow 
when analyzing the program, but traditional control-flow analysis collapses 
the state space into a finite space thus causes imprecise stack modeling.

To see how traditional CFA suffers from spurious return-flows, 
we can consider the following example:

@@

In traditional $k$-CFA algorithm and the AAM we showed in section \ref{aam}, 
the return-flow of invocation @(id 2)@
is going to both call-sites @(id 1)@ and @(id 2)@. Therefore, the
returned value @2@ for variable @y@ is also propagated to 
variable @x@ and then imprecise analysis result is arose.
This return-flow merging is inevitable even we increase the context-sensitivity.
If we use the monovariant analysis (i.e. 0CFA), the analyzing result would be @x@ and 
@y@ point to set @{1, 2}@, because the algorithm not distinguishes
@z@ comes from different call-sites.
Under $1$-CFA, the algorithm is able to distinguish that variable @z@ of function
@id@ has two different values at two call-sites, so variable @y@ 
would not be polluted by @1@. 
But two call-sites still share the same continuation and
it is still helplessness for separating the return-flows.
TODO: do need more details to explain this example? (as in P4F paper)

>>>>

As we mentioned previously, the components of states are lifted as arguments to
@aeval@. Therefore, the pattern matching on state also becomes just matching on
expression @e@.
\begin{lstlisting}
type Cont = Ans => Ans
def aeval(e: Expr, env: Env, store: BStore, time: Time, continue: Cont): Ans = {
  val new_time = (e::time).take(k)
  e match {
    case Let(x, App(f, ae), e) =>
      ......
    case ae if isAtomic(ae) => 
      ......
  }
}
\end{lstlisting}

@@

\textbf{First Version of Evaluation Function.}
Since our first version of @aeval@ just implements the non-deterministic
evaluation, and given the defined @nd@, the evaluation function @aeval@ can be easily
defined. For the case that @e@ is a @let@ expression,
the overall idea is to perform a two step depth-first evluation. \todo{refine}

\begin{lstlisting}
case Let(x, App(f, ae), e) =>
  val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
  nd[Clos, Ans](closures, Set[VS](), { case (clos, accouter, closnd) =>
    val Clos(Lam(v, body), c_env) = clos
    val baddr = allocBind(v, new_time)
    val new_env = c_env + (v -> baddr)
    val new_store = store.update(baddr, atomicEval(ae, env, store)) 
    aeval(body, new_env, new_store, new_time, (bodyvss: Set[VS]) => {
      nd[VS, Ans](bodyvss, Set[VS](), { case (vs, accinner, bdnd) => 
        val VS(vals, time, store) = vs
        val new_env = env + (x -> baddr)
        val new_store = store.update(baddr, vals)
        aeval(e, new_env, new_store, time, 
              /* accumulate the values of one path, and call next body value*/
              (evss: Ans) => bdnd(accinner ++ evss)) 
      },
      /* accumulate the values of pathes to accouter, and call next closure */
      (evss: Ans) => closnd(evss ++ accouter)) 
    })
  }, 
  continue)
\end{lstlisting}

The outer function call @nd@ evaluates over the set of possible closures of
@f@. For each closures, we go into its body with new environment and new store.
Remember that the function @aeval@ returns a set of @VS@ to its continuation,
so the continuation argument @bodyvss@ is a set of @VS@ that represents values
from multiple computation pathes when evaluating this single @body@ expression.

Then for the inner call of function @nd@ evaluates over the set of body values @bodyvss@.
For each @VS@, the time stamp and store in @VS@ will be instantiated to
@aeval@.
The inner application of @aeval@ returns set of values of one computation path to its
continuation, where the continuation will accumulate the result and transfer
evaluation to the next body value.

\begin{lstlisting}
case ae if isAtomic(ae) => 
  val ds = atomicEval(ae, env, store)
  continue(Set(VS(ds, new_time, store))) 
\end{lstlisting}

If the expression is atomic, then we just need to construct a new instance of
@VS@, return it to the @continue@ function.

@@

In the last section, we just established a computable pushdown control-flow
analysis through refunctionalization and caching. In this section,
we examine how pushdown control-flow is obtained and what we have done to
achieve it.

\textbf{The Problem of Return-flows.}
Pushdown control-flow is a property in analysis that precisely models
the runtime call-stack of the analyzed program. 
A pushdown control-flow analysis provides as exact as runtime return-flow 
when analyzing the program, but traditional control-flow analysis collapses 
the state space into a finite space thus causes imprecise stack modeling.

To see how traditional control-flow analysis suffers from spurious return-flows, 
we can consider the following example:

@@

In $k$-CFA algorithm or the abstract abstract machine we showed in section
\ref{aam},
The call sites @(id 2)@ and @(id 1)@ share the same return flow,
so the of invocation @(id 2)@ returns to both call-sites @[x (id 1)]@ and @[y (id 2)]@.
Therefore, the returned value @2@ for variable @y@ is also propagated to 
variable @x@ and then imprecise analysis result is arose.
This return-flow merging is inevitable even we increase the context-sensitivity.
If we use the monovariant analysis (i.e., 0CFA), the analyzing result would be @x@ and 
@y@ point to value set @{1, 2}@, because the algorithm not distinguishes
@z@ comes from different call-sites.
Under $1$-CFA, the algorithm is able to distinguish that variable @z@ of function
@id@ has two different values at two call-sites, so variable @y@ 
would not be polluted by @1@.
However, variable @x@ still points to value set @{1, 2}@.
Because two call-sites still share the same continuation and
it is still helplessness for separating the return-flows.

\textbf{Call/Return Matching through Refunctionalization and Caching.}
\todo{how refunc and caching help}
